\include{preamble}

\title{\textbf{% ECON 620\\
               Probability and Statistical Inference}}
\author{Tianqi Zhang\\
Emory University}
\date{Apr 17th 2025}

\begin{document}
\maketitle
\setcounter{tocdepth}{1} % Only show sections in the table of contents
\setcounter{section}{5}


\section{Measurable maps}
\subsection{Measurable maps}
\begin{df}{Measurable map}\\
Let \( (\Omega_1, \mathscr{F}_1) \), \( (\Omega_2, \mathscr{F}_2) \) be measurable spaces. \\
A map \( f: \Omega_1 \to \Omega_2 \) is measurable with respect to \( (\mathscr{F}_1, \mathscr{F}_2) \) if:
\[
f^{-1}(A) \in \mathscr{F}_1, \ \forall A \in \mathscr{F}_2.
\]
\noindent \textit{iff Pre-image of measurable sets are measurable.}
\end{df}

\begin{eg}{Indicator Function}\\
\noindent Consider \( \chi_A: (\Omega, \mathscr{F}) \to (\mathbb{R}, \mathscr{B}) \), where
\[
\chi_A(\omega) =
\begin{cases}
1, & \text{if } \omega \in A, \\
0, & \text{else}.
\end{cases}
\]
To show \( \chi_A \) is measurable, check all pre-images:
\[
\chi_A^{-1}(\emptyset) = \emptyset, \quad \chi_A^{-1}(\mathbb{R}) = \Omega, \quad \chi_A^{-1}(\{1\}) = A, \quad \chi_A^{-1}(\{0\}) = A^c.
\]
Since all are in \( \mathscr{F} \), \( \chi_A \) is measurable.

\end{eg}

\begin{eg}{}
\noindent Suppose \( f: (\Omega_1,\mathscr{F}_1)  \to (\Omega_2, \mathscr{F}_2) \) and \( g: (\Omega_2, \mathscr{F}_2) \to (\Omega_3, \mathscr{F}_3) \). If \( f \) and \( g \) are measurable, then \( g \circ f \) is measurable.

\noindent For any \( A \in \mathscr{F}_3 \),
\[
(g \circ f)^{-1}(A) = \underbrace{f^{-1}\underbrace{(g^{-1}(A)}_{\in \mathscr{F}_2})}_{\in\mathscr{F}_1}.
\]
Since \( g^{-1}(A) \in \mathscr{F}_2 \) and \( f^{-1} \) preserves measurability, \( (g \circ f)^{-1}(A) \in \mathscr{F}_1 \).
\end{eg}


\begin{prop}
\noindent Let \( (\Omega, \mathscr{F}) \), \( (\mathbb{R}, \mathscr{B}) \), and \( f, g: \Omega \to \mathbb{R} \) be measurable. Then:
\begin{enumerate}
    \item \( f+g, f-g \) are measurable,
    \item \( |f| \) is measurable.
\end{enumerate}	
\end{prop}

\begin{prop}{Measurable and Continuity}
\noindent Let \( f: X \to \mathbb{R} \) be a continuous function. Then \( f \) is measurable with respect to the Borel sigma algebra.
\end{prop}

\subsection{Probability measure to a random variable (Push-forward)}

\begin{df}{Push-forward measure}
Push-forward measure \( P_X \) is the measure induced on \( \mathbb{R}^d \) via the function \( X \), such that:
\[
\forall A \in \mathscr{B}^d, \; P_X(A) = P(X^{-1}(A)).
\]
\end{df}

\noindent For a probability (measure) space \((\Omega, \mathscr{F}, P)\) and a function (random variable) 
\[
X: (\Omega, \mathscr{F}, P) \to (\mathbb{R}^d, \mathscr{B}^d, \cdot),
\]
There are generally two ways to make $X$ measurable map: 
\begin{enumerate}
	\item Fix $\mathscr{B}^d$ and $\mathscr{F}$, ask if $X$ is measurable. 
	\item Alter $\mathscr{F}$ according to $X$ such that $X$ is measurable. 
\end{enumerate}

\subsubsection{Method 1: Validation}

Ask whether \( X \) is measurable with respect to the fixed \(\sigma\)-algebras:
\[
\forall A \in \mathscr{B}^d, \quad X^{-1}(A) \in \mathscr{F}.
\]
\noindent This directly checks the measurability condition. If satisfied, we say then \( X \) as a function is measurable.

\begin{itemize}
    \item \textit{Non-Measurability} shows by the \textit{same set having 2 different sizes (measures)}
\end{itemize}

\textbf{Criterion}
\textbf{Check (1):}
\begin{itemize}
    \item For \( \Omega \xrightarrow{X} \mathbb{R}^d \), \( \forall A \in \mathscr{B}^d \), \( X^{-1}(A) \in \mathscr{F} \), then \( X \) is measurable.
    \item \textcolor{red}{BUT: Too hard to validate \( \forall A \in \mathscr{B}^d \)}.
\end{itemize}

\subsubsection{Method 2: Generation}
Fix \(\mathscr{B}^d\), then construct a \(\sigma\)-algebra \(\mathscr{F}_X\) (generated by \(X\)) that ensures measurability:
\[
X^{-1}(\mathscr{B}^d) \coloneqq \{X^{-1}(A) \mid A \in \mathscr{B}^d\}.
\]
Define:
\[
\mathscr{F}_X \equiv \sigma(X) \quad \text{(the \(\sigma\)-algebra generated by \(X\))}.
\]
\noindent This is the smallest \(\sigma\)-algebra that makes \(X\) measurable:
\begin{itemize}
    \item \textit{Actually, \(\sigma(X)\) is the smallest \(\sigma\)-algebra for \(X\) to be measurable.}
    \item \(\sigma(X)\) contains all pre-images of \(\mathscr{B}^d\) under \(X\).
\end{itemize}

\begin{eg}{}\\
Let \( X(\omega) = \mathbb{I}_A(\omega) \) for some \( A \subseteq \Omega \), where \( \mathbb{I}_A \) is the indicator function of \( A \). Then the \(\sigma\)-algebra generated by \( X \) is:
\[
\sigma(X) = \{\emptyset, A, A^c, \Omega \}.
\]
This \(\sigma\)-algebra usually represents \textit{information sets} in practice.
\end{eg}

\begin{thm}{Measurability Validation}
\textbf{Solution:} It is sufficient to check the generators \( B \) of $\mathscr{B}$ such that \( X^{-1}(B)\subset \mathscr{F}\).\\
\end{thm}

\begin{eg}\\
 Let \( \mathcal{L} = \left\{ \prod_{i=1}^d (-\infty, x_i] \mid x_i \in \mathbb{R} \right\} \) be a generator of \( \mathscr{B}^d \).
 We have \( X^{-1}(\mathcal{L}) \subseteq \mathscr{F} \), then we generate the smallest \( \sigma \)-algebra:
    \[
    \sigma(X^{-1}(\mathcal{L})) \]
We need to show that
   \[\sigma(X) \subseteq \sigma(X^{-1}(\mathcal{L})) \subseteq \mathscr{F}.
    \]
which makes \( X \) measurable.
\end{eg}
\end{document}